---
title: "P8105_hw3_zj2358"
author: "Zhezheng Jin"
date: "2023-10-10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE) 
library(p8105.datasets)
library(tidyverse)
library(knitr)
```

## Problem 1

```{r}
data("instacart")
```


Description of the dataset:

The dataset `instacart` has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Variables in the dataset include `r colnames(instacart)`.
Some key variables include unique order identifiers (`order_id`, e.g., 13749), the sequence products are added to the cart (`add_to_cart_order`, e.g., 2 for the second item), indicators for product reorders (`reordered`, e.g., 0 for previously ordered), the day an order was placed (`order_dow`, e.g., 2 for Tuesday), and specific product names (`product_name`, e.g., Clean Linen Candle).


How many aisles are there, and which aisles are the most items ordered from?

```{r}
num_aisles <- n_distinct(instacart$aisle)

top_aisle <- instacart %>%
  group_by(aisle) %>%
  summarise(count = n()) %>%
  arrange(-count) %>%
  slice(1) %>%
  pull(aisle) 
```

There are `r num_aisles` unique aisles, and the most items are ordered from the `r top_aisle`.


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_plot <- instacart %>%
  group_by(aisle) %>%
  summarise(item_count = n()) %>%
  filter(item_count > 10000) %>%
  arrange(-item_count)
ggplot(aisle_plot, aes(x = reorder(aisle, item_count), y = item_count)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  labs(title = "Number of Items Ordered in Each Aisle",
       x = "Aisle",
       y = "Number of Items ordered") +
  theme_minimal()
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
popular_items <- instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(item_count = n(), .groups = "drop") %>%
  arrange(aisle, -item_count) %>%
  group_by(aisle) %>%
  slice_head(n = 3)

knitr::kable(popular_items,caption = "Top 3 Items in Selected Aisles", align = c('l', 'l', 'r'))
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
product_hour <- instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day, na.rm = TRUE), .groups = "drop") %>%
  spread(key = order_dow, value = mean_hour) %>%
  select(product_name, `0`, `1`, `2`, `3`, `4`, `5`, `6`) 

knitr::kable(product_hour, caption = "Mean Hour of Order by Product and Day of the Week", digits = 2)
```

## Problem 2

```{r}
data("brfss_smart2010")
```

Data cleaning

```{r}
colnames(brfss_smart2010)
brfss_smart2010 <- brfss_smart2010 %>%
  rename(
    location_abbr = Locationabbr,
    location_desc = Locationdesc,
    class_id = ClassId,
    data_source = DataSource,
    topic_id = TopicId,
    location_id = LocationID,
    question_id = QuestionID,
    resp_id = RESPID,
    geo_location = GeoLocation
  ) %>%
    janitor::clean_names() %>%
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                           ordered = TRUE))
```

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
states_with_7_or_more_locations <- brfss_smart2010 %>%
  filter(year %in% c(2002, 2010)) %>%
  group_by(year, location_abbr) %>%
  summarise(num_entries = n_distinct(location_desc)) %>%
  filter(num_entries >= 7)

states_2002 <- states_with_7_or_more_locations %>%
  filter(year == 2002)  %>%
  pull(location_abbr)

states_2010 <- states_with_7_or_more_locations %>%
  filter(year == 2010) %>%
  pull(location_abbr)
```

In 2002, states observed at 7 or more locations were `r paste(states_2002, collapse=", ")`. In 2010, states observed at 7 or more locations were `r paste(states_2010, collapse=", ")`. 


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r}
excellent_responses <- brfss_smart2010 %>%
  filter(response == "Excellent") %>%
  group_by(year, location_abbr) %>%
  summarise(avg_data_value = round(mean(data_value, na.rm = TRUE),2))
```

Make a “spaghetti” plot of this average value over time within a state 

```{r}
ggplot(excellent_responses, aes(x = year, y = avg_data_value, group = location_abbr)) +
  geom_line(aes(color = location_abbr), alpha = 0.7) +
  labs(title = "Average value trends for each state",
       y = "Average Data Value") +
  theme_minimal() 
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
ny <- brfss_smart2010 %>%
  filter(year %in% c(2006, 2010) & location_abbr == "NY")

ggplot(ny, aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_wrap(~ year, ncol = 2) +
  labs(title = "Distribution of Data Value for responses in NY (2006 & 2010)",
       y = "Data Value",
       x = "Response") +
  theme_minimal()
```

## Problem 3

Load and clean the datasets

```{r}
nhanes_accel <- read_csv("nhanes_accel.csv")  %>%
  janitor::clean_names()
nhanes_covar <- read_csv("nhanes_covar.csv") %>%
  janitor::clean_names()
```

Tidy the data

```{r}
nhanes_covar <- nhanes_covar[-c(1:4),] 
colnames(nhanes_covar) <- c("seqn", "sex", "age", "bmi", "education")
nhanes_covar$seqn <- as.integer(nhanes_covar$seqn)
nhanes_covar$age <- as.numeric(nhanes_covar$age)
nhanes_covar$bmi <- as.numeric(nhanes_covar$bmi)

nhanes_accel$seqn <- as.integer(nhanes_accel$seqn)
```

Filter the data

```{r}
nhanes_covar <- nhanes_covar %>% 
  filter(age >= 21, !is.na(sex), !is.na(age), !is.na(bmi), !is.na(education))
```

Merge 

```{r}
nhanes <- inner_join(nhanes_covar, nhanes_accel, by = "seqn")
```

Encode data

```{r}
nhanes$sex <- factor(nhanes$sex, levels = c("1", "2"), 
                          labels = c("Male", "Female"))

nhanes$education <- factor(nhanes$education, 
                               levels = c("1", "2", "3"), 
                               labels = c("Less than high school", 
                                          "High school equivalent", 
                                          "More than high school"))
```

Produce a Reader-Friendly Table

```{r}
education_table <- table(nhanes$sex, nhanes$education)
knitr::kable(education_table, 
             caption = "Number of Men and Women in Each Education Category")
```

For the Table: Examine the counts for each gender within the education categories. Are there more males or females in specific education levels? Is there a noticeable difference?


Visualization of the age distributions

```{r}
ggplot(nhanes, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distribution by Sex and Education",
       x = "Education Level",
       y = "Age",
       fill = "Sex") +
  theme_minimal()
```

For the Visualization:

Look at the median age (the horizontal line in the middle of the box) for each category. Which group is younger or older on average?
Observe the spread (size of the box). Which group has a wider age range?
Are there any outliers (points outside the "whiskers" of the boxplot)?




Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.


